Assessing Digital Elevation Models with different image overlap using photogrammetry

### 2021-04-24 Paper.tex
20200424 backup

% KAO: Use times symbol
\abstract{
Photogrammetry is easy to use these days and can be used to create a digital elevation model of an area. 
Image overlap plays an important role in the process of photogrammetry.
A high overlap takes more time to gather in the field and takes longer to process.
Less overlap usually results in a less accurate calculations. 
In this paper the overlap of an area is artificially reduced and digital elevation models have been created with different image overlaps.
The digital elevation models are compared to see how reducing the image overlap influences the result.
}

@@

\subsection{Study Area}\label{sec:Study Area}
The data covers an area in Ghana, about 20km south east of Ejura (lat. 7.3064212, lon. -1.164493). 
The area of interest is about 25000 m\textsuperscript{2}. 
The area consists mostly of sanvannah.
There are a few trees present.
\begin{figure}[h]
    \centering
    \includegraphics[width=4cm]{locationwide.png}
    \includegraphics[width=4cm]{imgloc.png}
    \caption{Area of Interest and Unprocessed Image GPS Locations}
    \label{fig:areaofinterest}
\end{figure}
Each image has a resolution of 4000 times 3000 and was taken by a DJI Phantom 3.
The images all contain GNNS location (WGS84). 
The complete data set consists of 111 images.
Their exact location can be found in figure \ref{fig:areaofinterest}.

\subsection{Image Processing}
For image processing and DEM generation Agisoft Professional edition 1.6.5 (Agisoft LLC., St. Peterburg, Russia) was used \citep{AgisoftMetashape}.
All data have been processed on the cloud service provided by Agisoft (\url{https://cloud.agisoft.com}).
The cloud processing is done on a vCPU (Intel Xeon E5 2686 v4) with two NVIDIA Tesla M60 GPU's and 240 GB of RAM.

First, the complete data set has been aligned on medium quality.
Next, the dense cloud was built on medium quality (no depth filtering), resulting in a dense cloud of 3,268,264 points.
A mesh was built using the the triangulated irregular network (TIN) algorithm by \citet{axelsson1999processing}. 
After the mesh the Reduce Overlap function was used to reduce the amount images used in the process of photogrammetry. 
The mesh was set to create a 2.5D height field.

\textbf{Reduce Overlap.} 
The algorithm of the Reduce Overlap function tries to select a minimal amount of images such that each point of the model is observed from N locations/angles.
If a point is observed multiple times from the same location/angle, than this still counts as one. 
The exact algorithm and parameters behind the Agisoft Reduce Overlap function are unknown.
The input dataset was a 2.D height field mesh. 
Because in this paper the comparison will be made with a DEM, a height field was computed.
In Agisoft Metashape the 'Reduce Overlap' had three different settings, low, medium and high. 
All of the settings are used to compute the images for further processing.

\textbf{Camera Locations.}
In figure \ref{fig:cameralocation} the estimated camera locations can be found. The camera selection is done by the Reduce Overlap algorithm from Agisoft Metashape. The estimated camera location is done in the image processing alogrithm of Metashape.
The images are matched based on an adapted version of the SIFT algorithm \citep{lowe1999object, AgisoftMetashape}.

\begin{figure}[h]
    \centering
    \includegraphics[width=4cm]{loc_low.png}
    \includegraphics[width=4cm]{loc_med.png}
    \includegraphics[width=4cm]{loc_full.png}
    \includegraphics[width=4cm]{loc_high.png}
    \caption{Estimated camera location. 
    White: used in the image processing, red: disabled in the image processing.\\
    Clockwise (starting left top): low, medium, high and full settings}
    \label{fig:cameralocation}
\end{figure}

@@

\subsection{DEM}
The source data for the DEM's are the dense cloud results. 
The Dense Cloud contains about the same amount of points with the exception for the lowest Reduce Overlap setting.
This setting has resulted in the dense cloud with the most points.
The TIN algorithm is used to built the DEM's \citep{axelsson1999processing}.
The image overlap is now changed, therefore the DEM's have a slightly different extent.
Because of this the largest common extent of the DEM's will be used to crop the DEM's resulting in four DEM's with an equal extent.
The DEM's will be exported from Agisoft Metashape with the same pixel resolution of 17.5 cm per pixel.
The DEM's will also be scaled to check the differences between the DEM's without looking at the absolute values.
The scaling will be done by subtracting the mean and dividing by the standard deviation.
For further comparison all of the DEM's will be subtracted from the full model DEM.
This will result in three new DEM's that contain differences instead of height values.

@@

In figure \ref{fig:ViolinPlot} a violin plot containing the difference of the DEM between the different Reduce Overlap settings can be found.

\begin{figure}[h]
    \centering
    \includegraphics[width=8cm]{ViolinPlot2.png}
    \caption{Violin plot of the difference between the full DEM and the reduced DEM's}
    \label{fig:ViolinPlot}
\end{figure}

@@

There is a large different between the absolute values between the different DEM's, this can be seen visually in figure \ref{fig:DemPlot_unscaled} and also when comparing the boxplot values in figure \ref{fig:BoxPlot_unscaled}.
With the use of at least three ground control points the error in the absolute values of the can be reduced \citep{AssessingUAVGCPS, GCPbetterAccuracy}.

\section{Conclusion}
Using different image overlap in the same area results in DEM's with different absolute values.
The two image dataset with the least images show the largest difference in absolute value range compared to the original dataset with all images. 
The image dataset with the highest image count, has a DEM that is most similar compared to the original dataset.
When the scaled values are compared there seems to be a minimal difference between the DEM's

>>>>

% KAO: Use times symbol
\abstract{
Photogrammetry is easy to use these days and can be used to create a digital elevation model of an area. 
Image overlap plays an important role in the process of photogrammetry.
A high overlap takes more time to gather in the field and takes longer to process.
Less overlap usually results in a less accurate calculations. 
In this paper the overlap of an area is artificially reduced and digital elevation models have been created with different image overlaps.
The digital elevation models are compared to see how reducing the image overlap influences the result. 
A boxplot has been made to compare to the absolute values of the DEM's.
Large difference are visible, however when the scaled DEM's are plotted in a boxplot the difference appear to be minimal.
Finally a violin plot has been made of the complete image set DEM minus the reduces image set DEM's.
The more images are used, the more the difference of the DEM's is centered around 0.
}

@@

\subsection{Study Area}\label{sec:Study Area}
The data covers an area in Ghana, about 20km south east of Ejura (lat. 7.3064212, lon. -1.164493). 
The area of interest is about 25000 m\textsuperscript{2}. 
The area consists mostly of savanna area.
There are a few trees present on the observed location.
\begin{figure}[h]
    \centering
    \includegraphics[width=4cm]{locationwide.png}
    \includegraphics[width=4cm]{imgloc.png}
    \caption{Area of Interest and Unprocessed Image GPS Locations}
    \label{fig:areaofinterest}
\end{figure}
Each image has a resolution of 4000 times 3000 and was taken by a DJI Phantom 3.
The images all contain GNNS location (WGS84). 
The complete data set consists of 111 images.
Their exact location can be found in figure \ref{fig:areaofinterest}.

\subsection{Image Processing}
For image processing and DEM generation Agisoft Professional edition 1.6.5 (Agisoft LLC., St. Peterburg, Russia) was used \citep{AgisoftMetashape}.
All data have been processed on the cloud service provided by Agisoft (\url{https://cloud.agisoft.com}).
The cloud processing is done on 32 vCPU's (Intel Xeon E5 2686 v4) with two NVIDIA Tesla M60 GPU's and 240 GB of RAM.

First, the complete data set has been aligned on medium quality.
Next, the dense cloud was built on medium quality (no depth filtering), resulting in a dense cloud of 3,268,264 points.
A mesh was built using the the triangulated irregular network (TIN) algorithm by \citet{axelsson1999processing}. 
finally after the mesh the Reduce Overlap function was used to reduce the amount images used in the process of photogrammetry. 
The mesh was set to create a 2.5D height field.

\textbf{Reduce Overlap.} 
The algorithm of the Reduce Overlap function tries to select a minimal amount of images such that each point of the model is observed from N locations/angles.
If a point is observed multiple times from the same location/angle, than this still counts as one. 
The exact algorithm and parameters behind the Agisoft Reduce Overlap function are unknown.
The input dataset was a 2.D height field mesh. 
Because in this paper the comparison will be made with a DEM, a height field was computed.
In Agisoft Metashape the 'Reduce Overlap' had three different settings, low, medium and high. 
All of the settings are used and the remaining images are used for further processing.
In \ref{tab:ImageProcessing} the amount of images for each Reduce Overlap setting can be found.

\textbf{Camera Locations.}
In figure \ref{fig:cameralocation} the estimated camera locations can be found. The estimated camera location is estimated by the image processing alogrithm of Metashape.
The images are matched based on an adapted version of the SIFT algorithm \citep{lowe1999object, AgisoftMetashape}.

\begin{figure}[h]
    \centering
    \includegraphics[width=1.9cm]{loc_low.png}
    \includegraphics[width=1.9cm]{loc_med.png}
    \includegraphics[width=1.9cm]{loc_high.png}
    \includegraphics[width=1.9cm]{loc_full.png}
    \caption{Estimated camera location. 
    White: used in the image processing, red: disabled in the image processing.\\
    Clockwise (starting left top): low, medium, high and full settings}
    \label{fig:cameralocation}
\end{figure}

@@

\textbf{DEM.}
The source data for the DEM's are the dense cloud results. 
The Dense Cloud contains about the same amount of points with the exception for the lowest Reduce Overlap setting.
This setting has resulted in the dense cloud with the most points.
The TIN algorithm is used to built the DEM's \citep{axelsson1999processing}.
The image overlap is now changed, therefore the DEM's have a slightly different extent.
Because of this the largest common extent of the DEM's will be used to crop the DEM's resulting in four DEM's with an equal extent.
The DEM's will be exported from Agisoft Metashape with the same pixel resolution of 17.5 cm per pixel.
The DEM's will also be scaled to check the differences between the DEM's without looking at the absolute values.
The scaling will be done by subtracting the mean and dividing by the standard deviation.
For further comparison all of the DEM's will be subtracted from the full model DEM.
This will result in three new DEM's that contain differences instead of height values.

@@

In figure \ref{fig:ViolinPlot} a violin plot containing the difference of the DEM between the different Reduce Overlap settings can be found. 
It is visible that the overall error of the scaled DEM's varies from around -5.0 to 5.0.
For the high settings, the differences are more centered around zero, while this is not the case for low and medium settings.
For medium setting the difference is slightly more centered around 0.0 than low settings.

\begin{figure}[h]
    \centering
    \includegraphics[width=8cm]{ViolinPlotOrdered.png}
    \caption{Violin plot of the difference between the full DEM and the reduced DEM's}
    \label{fig:ViolinPlot}
\end{figure}

@@

There is a large different between the absolute values between the different DEM's, this can be seen visually in figure \ref{fig:DemPlot_unscaled} and also when comparing the boxplot values in figure \ref{fig:BoxPlot_unscaled}.
With the use of at least three ground control points the error in the absolute values of the can be reduced \citep{AssessingUAVGCPS, GCPbetterAccuracy}.
Using GCP's in the different image dataset will reduce the absolute error, further research is needed how much is influence the relative error of the different DEM's

\section{Conclusion}
Using different image overlap in the same area results in DEM's with different absolute values.
The two image dataset with the least images show the largest difference in absolute value range compared to the original dataset with all images. 
The image dataset with the highest image count, has a DEM that is most similar compared to the original dataset.
When the scaled values are compared there seems to be a minimal difference between the DEM's.
If more images are used in the dataset, the difference between the DEM's compared to the full images set becomes smaller; the DEM differences are more centered around 0.
